Reasoning About a Machine with Local Capabilities

### 2018-05-08 toplas/paper.tex
Added first draft for presentation illustrations (caption and references missing!)

% \newtheorem{theorem}{Theorem}
% \newtheorem{lemma}{Lemma}

@@

\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.

>>>>

% \newtheorem{theorem}{Theorem}
% \newtheorem{lemma}{Lemma}
\input{preamble-illu}

@@

\begin{figure}[htb]
  \centering
  \begin{subfigure}{0.32\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \actsf{(0,5)}{(4.5,7)}{Our stack frame}

    % Unused part
      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=red] {\footnotesize ostk};
      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=red] {\footnotesize opc};
      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=red] {\footnotesize ...};
      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=red] {\footnotesize ...};

    % Unused part
      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=teal] {\footnotesize 0};
      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=teal] {\footnotesize 0};
      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=teal] {\footnotesize 0};
      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=teal] {\footnotesize 0};

    % Unused part
      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}

  \begin{subfigure}{0.32\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \inactsf{(0,5)}{(4.5,7)}{Our stack frame}

    % second attacker
      \actadv{(0,7)}{(4.5,10)}{Callee\\\footnotesize stack frame}

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5] {\footnotesize 0};
      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5] {\footnotesize 0};
      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5] {\footnotesize 0};
      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5] {\footnotesize 0};

    % Unknown part
      \node[color=red] at (2.25,9) {\huge ?};

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker ..
      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}

    % trusted stack frame
      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
      \foreach \x in {5,5.5,...,10}
      {
        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}

  \caption{test}
  \label{fig:ret-adv}
\end{figure}
\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.
